---
output: html_document
---

# datenguideR 

```{r, include = FALSE}
## install the following packages:
# devtools::install_github("CorrelAid/datenguideR")
if(!require("pacman")) install.packages("pacman")


pacman::p_load(tidyverse, ggplot2, ggthemes, knitr, datenguideR)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4,
  out.width = "100%"
)


```


Access and download German regional statistics from Datenguide <http://datengui.de>. `datenguideR` provides a wrapper for their GraphQL API and also includes metadata for all available statistics and regions.


## Usage

First, install datenguideR from GitHub:

```{r, eval=F}
devtools::install_github("CorrelAid/datenguideR")
```

Load package:

```{r}
library(datenguideR)
```


## Examples

Get IDs of all available NUTS-1 regions:

```{r}
datenguideR::dg_regions %>%
  dplyr::filter(level == "nuts1") %>%
  knitr::kable()
```

Get all available meta data on statistics, substatistics, and parameters:

```{r}
datenguideR::dg_descriptions
```

### `dg_search`

You can also use `dg_search` to look for a variable of interest. The function will match your string with any strings in the `dg_descriptions` data frame, returning only rows with those matches.

Looking for variables where the string *"vote"* appears somewhere in the documentation:

```{r}
dg_search("vote")
```

Note: Descriptions of variables are also available in English now! Translated via the [`googleLanguageR`](https://github.com/ropensci/googleLanguageR) package.

```{r}
dg_search("vote") %>% 
  dplyr::select(stat_name, dplyr::contains("_en"))
```

### `dg_call`

The main function of the package is `dg_call`. It gives access to all API endpoints.

Simply pick a statistic and put it into `dg_call()` (infos can be retrieved from `dg_descriptions`).

For example:

 - `stat_name`: AI0506 *(Wahlbeteiligung, Bundestagswahl)*
 - `region_id`: 11 (stands for Berlin)

```{r}
dg_call(region_id = "11",
        year = 2017,
        stat_name = "AI0506")



```



A slightly more complex call with substatistics:

  - `stat_name`: BETR08 *(Landwirtschaftliche Betriebe mit Tierhaltung)*
  - `substat_name`: TIERA8 *(Landwirtschaftliche Betriebe mit Viehhaltung)*
  - `parameter`:
    - TIERART2 *(Rinder)*
    - TIERART3 *(Schweine)*
    
```{r}

dg_call(region_id = "11", 
        year = c(2001, 2003, 2007), 
        stat_name = "BETR08", 
        substat_name = "TIERA8", 
        parameter = c("TIERART2", "TIERART3")) 
```

If you give no parameters for a substat, it will default to return results for all of them.

```{r}
dg_call(region_id = "11", 
        year = c(2001, 2003, 2007), 
        stat_name =  "BETR08", 
        substat_name = "TIERA8") 
```


### Multiple regions with `nuts_nr` or `lau_nr`

Instead of specifying a `region_id` for individual Bundesland (state) data you can also use `nuts_nr` to receive data for *NUTS-1*, *NUTS-2* and *NUTS-3*.

Just leave `region_id` blank and provide either a `nuts_nr` (or `lau_nr`) to get data for multiple regions at once.

 - `stat_name`: AI0506 *(Wahlbeteiligung, Bundestagswahl)*

```{r}
dg_call(nuts_nr = 1,
        year = 2017,
        stat_name = "AI0506")



```

  - `stat_name`: BETR08 *(Landwirtschaftliche Betriebe mit Tierhaltung)*
  - `substat_name`: TIERA8 *(Landwirtschaftliche Betriebe mit Viehhaltung)*

```{r}
dg_call(nuts_nr = 1,
        year = c(2001, 2003, 2007), 
        stat_name = "BETR08", 
        substat_name = "TIERA8") 
```

  - `stat_name`: BAU018 *Total non-residential buildings*
  - `substat_name`: BAUAHZ *Main Type of Heating*

	
 

```{r}
dg_call(nuts_nr = 1,
        stat_name =  "BAU018",
        substat_name = "BAUAHZ",
        year = 2016)
```



### `dg_map`

You can also use `dg_map` to plot retrieved data on a Germany map. This currently only supports NUTS-1 and NUTS-2. Arguments are (mostly) identical to `dg_call`.


```{r}
dg_map(nuts_nr = 1,
        year = 2017,
        stat_name = "AI0506")

```

The output is a ggplot object and can be manipulated further.

```{r}
dg_map(nuts_nr = 1,
        year = 2017,
        stat_name = "AI0506") +
  ggthemes::theme_map() +
  ggplot2::scale_fill_viridis_c("Voter Turnout") +
  ggplot2::ggtitle("Voter Turnout in German Parliamentary Election (2017)") +
  ggplot2::theme(legend.position = "right")

```

You can also return the data and use your own plotting functions with `return_data = TRUE`.

```{r}
turnout_dat <- dg_map(nuts_nr = 1,
        year = 2017,
        stat_name = "AI0506",
        return_data = T)

turnout_dat

```



